---
title: "An easy day"
output:
  html_document:
    toc: true
---
<style>
div.hideme {display: none}
</style>
<script>
doclick=function(e){
        code = e.nextSibling.nextSibling
        if(code.style.display=="block") {
                 code.style.display='none';
                 e.textContent="Show Code"
        }else{
                 code.style.display="block";
                 e.textContent="Hide Code"
        }
}
</script>

```{r global_options, echo=FALSE, message=FALSE}
source("readlogs.R")
source("bot_collector.R")
source("geo_ip.R")

require(data.table)
require(reshape2)
```


```{r load_data, cache=TRUE, message=FALSE, results='hide', echo=FALSE}
yesterday <- Sys.Date() - 1
yesterfor <- format(yesterday, "%A %d-%m-%Y")
dt.log <- remote.data.table(yesterday)
dt.log[, agent:=ip.agent(ip_address)]
dt.events <- data.table(read.csv("ref/stats_events.csv", stringsAsFactors = FALSE))
setkey(dt.events, "id")
```
*Report based on the log file from `r yesterfor` (`r nrow(dt.log)` events)* 

### Geographic spread of client IP-addresses   
<!--<button class="hidecode" onclick="doclick(this);">Show Code</button>
<div class="hideme">-->
```{r geo_spread, results='hide', message=FALSE, echo=FALSE}
## get the unique IP-addresses
dt.ips <- dt.log[, j=list(frequency = .N), by = c("ip_address")]
dt.ips[, agent:=ip.agent(ip_address)]
setkey(dt.ips, "ip_address")
# 
## get a table with geo information on the addresses
dt.gip <- geo.info(dt.ips$ip_address)
dt.ll <- dt.gip[!(is.na(longitude) | is.na(latitude)),]
dt.ll <- merge(dt.ll, dt.ips, by = "ip_address")

require(maps)
require(mapdata)

# do some colouring
agents <- levels(factor(dt.ll$agent))
cols <- rainbow(length(agents), start=0.3, alpha=0.3)
sats <- rainbow(length(agents), start=0.3, alpha=1)
dt.col <- data.table(agent = agents, color = cols, sat = sats)
dt.world <- merge(dt.ll, dt.col, by = "agent")
max.color <- dt.world[which.max(frequency),]$sat
max.agent <- dt.world[which.max(frequency),]$agent
max.frequency <- dt.world[which.max(frequency),]$frequency
```
<!--</div>-->
```{r plot_geo_spread, fig.width=9, fig.height=6, echo = FALSE}
op <- par(mar = c(0, 1, 4, 1))
map('worldHires', , mar = c(0,0,2,0))
points(x = dt.world$longitude, y = dt.world$latitude, col=dt.world$color, 
       pch=19, cex = log10(dt.world$frequency) * 3)
# legend point
rect(xleft = -180, xright = -50, ybottom = -90, ytop = -60, col = "white", border = "white")
points(x = -160, y = -70, pch=21, cex = log10(1000) * 3, 
       col = max.color, bg="white")
points(x = -160, y = -70, pch=21, cex = log10(100) * 3, 
       col = max.color, bg="white")
points(x = -160, y = -70, pch=21, cex = log10(10) * 3, 
       col = max.color, bg="white")
points(x = -160, y = -70, pch=21, cex = log10(2) * 3, 
       col = max.color, bg="white")

text(x = -147, y = -80, labels="number of requests per IP-address:\n1000, 100, 10, 2 requests", 
     cex = 0.8, pos = 4)

legend(x = "bottomright", legend = dt.col$agent, col=dt.col$sat, 
       pch = 19, cex = 0.8, horiz=TRUE, bg = "white")
title(main = paste("Spread of requests, world-wide, ", yesterfor))
par(op)
```

The number of requests per IP-address world-wide is shown in the
graphic above. Each circle represents one client IP, the diameter of the
circle corresponds to the number of requests from that client on a logarithmic scale.
Color codes show the type of client we are dealing with. Usually a large
amount of search bots is active and concentrated on the west-coast of
the United States. IP-addresses from private ranges are not geo-tracked.

The top-client of the day was a `r max.agent` IP-address, with `r max.frequency` requests.

<table><tr><td width="63%">
```{r plot_eu_spread, echo = FALSE, fig.height=4.3, fig.width=5.8}
#xl <- -11; xr <- 25; yb <- 45; yt <- 58
xl <- -11; xr <- 29; yb <- 42; yt <- 58.5
dt.eu <- dt.world[longitude >= xl & longitude <= xr & latitude >= yb & latitude <= yt]

op <- par(mar = c(1, 1, 4, 1))
map('worldHires', xlim=c(xl, xr), ylim=c(yb, yt), mar = c(0,0,3,2), fill=TRUE, col="lightyellow")
points(x = dt.eu$longitude, y = dt.eu$latitude, col=dt.eu$color, 
       pch=19, cex = log10(dt.eu$frequency) * 3)
title(main = paste("Spread of requests, neighbouring countries"),
      cex.main = 0.95)
par(op)
```
</td><td width="37%">
```{r plot_nl_spread, echo=FALSE, fig.height=4.3, fig.width=3.4}
## nl_spread
xl <- 3.3; xr <- 7.25; yb <- 50.75; yt <- 53.49
dt.nl <- dt.world[longitude >= xl & longitude <= xr & latitude >= yb & latitude <= yt]

op <- par(mar = c(1, 1, 4, 1))
map('worldHires', xlim=c(xl, xr), ylim=c(yb, yt), mar = c(0,2,3,0))
points(x = dt.nl$longitude, y = dt.nl$latitude, col=dt.nl$color, 
       pch=19, cex = log10(dt.nl$frequency) * 3)
title(main = paste0("Spread of requests\nNetherlands"), 
      cex.main = 0.95)
par(op)
```
</td></tr></table>

### Summary of events
 
<button class="hidecode" onclick="doclick(this);">Show Code</button>
<div class="hideme">
```{r event_frequency, results='hide'}
## counts per event_id, agent
dt.fat <- dt.log[, j=list(frequency = .N), by = c("event_id", "agent")]
dt.fat.wide <- dcast(dt.fat, event_id ~ agent, value.var = "frequency", fill = "")
setnames(dt.fat.wide, old = "event_id", new = "id")

## counts per event
dt.counts <- dt.log[, j=list(frequency = .N), by = c("event_id")]
setnames(dt.counts, old = "event_id", new = "id")

## merges with event description and split-counts per agent type
dt.counts <- merge(dt.counts, dt.events[, c(1, 4), with = FALSE], by = "id", all.x = TRUE)
dt.counts <- merge(dt.counts, dt.fat.wide, by = "id", all.x = TRUE)
align <- rep("r", ncol(dt.counts) - 1)
align[2] <- "l"
```
</div>

`r knitr::kable(dt.counts[order(frequency, decreasing = TRUE), -1, with = FALSE], 
format = "markdown", align = align)`

The table above shows a summary of the registered events during `r yesterfor`. 
The column **frequency** gives the total count of the events in **event description**.
The column **normal** gives the count per event that was triggered by a 'normal'
IP-addresses, the column(s) **private.x** give the score from in-house IP-addresses.
The rest of the requests were triggered by various search bots.

*****************
<small>
Information based on `r paste0("statistics.log.", yesterday)`.   
Search bots detection was based on information from http://www.iplists.com/nw/ and
http://chceme.info/ips/ (there are more bots than detected).   
Geograhic information on IP-addresses from http://www.telize.com    
Report generated: `r format(Sys.time(), "%A %d-%m-%Y %H:%M:%S")`   
</small>

<small>Logged events represent only a portion of user activity or requests to the server.
An *event* is always triggered by a client request; some requests trigger more than
one event. When speaking of *requests* above, we actually mean *events*.
</small>

*****************



